---
title: "Symlink Tool Intro Vignette"
output: 
  html_document:
    toc: true
    toc_float: true
# This is lighter weight, but doesn't support TOC
# output: rmarkdown::html_vignette
# vignette: >
#   %\VignetteIndexEntry{symlink_tool_vignette}
#   %\VignetteEngine{knitr::rmarkdown}
#   %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

# Chunk options

# This is a nice default so log output doesn't wrap
options(width = 320)

# save the built-in output hook
hook_output <- knitr::knit_hooks$get("output")

# set a new output hook to truncate text output
# if the output is too long, it will be truncated like:
# 
# top output
# ...
# bottom output
knitr::knit_hooks$set(output = function(x, options) {
  if (!is.null(n <- options$out.lines)) {
    x <- xfun::split_lines(x)
    if (length(x) > n) {
      # truncate the output
      # x <- c(head(x, n), "....\n")
      x <- c(head(x, n/2), '....', tail(x, n/2 + 1))
    }
    x <- paste(x, collapse = "\n")
  }
  hook_output(x, options)
})
```

```{r setup, include = FALSE}
# Installing locally only allows main branch - switching to load_all for development
devtools::load_all()
# library(vmTools, lib.loc = file.path("/mnt/share/code", Sys.info()[["user"]], "r_pkgs"))
library(data.table)
library(fs)
```

```{r utils, include = FALSE}

# Defining a couple vignette utilities

# Print the folder output 'tree' structure
print_tree <- function() {fs::dir_tree(root_base, recurse = TRUE)}

# Make a directory with desired defaults without cluttering the vignette
make_directory <- function(path){
  dir.create(path, recursive = TRUE, showWarnings = FALSE)
}

# print a symlink's target from the file system 
print_symlink <- function(symlink_type){
  print(grep(symlink_type, system(paste("ls -alt", root_input), intern = TRUE), value = TRUE))
}

#' Get output directory for results to save in.
#'
#' Returns a path to save results in of the form "YYYY_MM_DD.VV".
#'
#' @param root path to root of output results
#' @param date character date in form of "YYYY_MM_DD" or "today". "today" will be interpreted as today's date.
get_output_dir <- function(root, date) {
  if (date == "today") {
    date <- format(Sys.Date(), "%Y_%m_%d")
  }
  cur.version <- get_latest_output_date_index(root, date = date)
  
  dir.name <- sprintf("%s.%02i", date, cur.version + 1)
  return(dir.name)
}

#' get the latest index for given an output dir and a date
#'
#' directories are assumed to be named in YYYY_MM_DD.VV format with sane
#' year/month/date/version values.
#'
#' @param dir path to directory with versioned dirs
#' @param date character in be YYYY_MM_DD format
#'
#' @return largest version in directory tree or 0 if there are no version OR
#' the directory tree does not exist
get_latest_output_date_index <- function(root, date) {
  currentfolders <- list.files(root)
  
  # subset to date
  pat <- sprintf("^%s[.]\\d{2}$", date)
  date_dirs <- grep(pat, currentfolders, value = TRUE)
  
  if (length(date_dirs) == 0) {
    return(0)
  }
  
  # get the index after day
  date_list <- strsplit(date_dirs, "[.]")
  
  inds <- unlist(lapply(date_list, function(x) x[2]))
  if (is.na(max(inds, na.rm = TRUE))) inds <- 0
  
  return(max(as.numeric(inds)))
}
```




# What is the SymLink Tool?



## Assumptions

This tool assumes you have a large set of output folders for runs of your pipeline, and you store them on the cluster file system.

If you're already using a central database to manage your output versions, you probably don't need this.  You can stop reading now!




## Tell Me All About It


The SymLink Tool is a way for researchers to manage data pipeline output runs.  It will:

1. Create a 'best' symlink to the 'best' pipeline run of your outputs.
    1. This is just a shortcut link in the `H:` drive i.e. on the Cluster File System.
    1. It allows you to have a **stable** file path to results **that you and others don't need to update**.
    1. Safely update the 'best' symlink when you, the researcher, want to 'promote' a new pipeline run to being 'best'.



> OK, that's nice, but symlinks are pretty easy to make.  What else do I (the researcher) get **for free** if I use this tool?


The most important thing you get are some simple **logs** that automatically keep track of which folders have been 'best.'

1. Maintain a **central log.**
    1. Which pipeline run is currently 'best'?
    1. Which older runs were _ever_ 'best' in case you need to 'roll back' what you consider 'best'.
    1. Track deletion of old pipeline runs.
1. Maintain **run-specific logs.**
    1. When was this runs started?  (the file system actually _doesn't_ keep track of this!)
    1. When was it marked 'best'?
    1. If it stopped being 'best', when did that happen?


> Hmm, is that all?  I feel like I could keep an excel document or HUB page that does the same thing?


That's very true.  But maybe there are pipeline runs you want to track for different reasons.  

- What if you want to keep a pipeline run for a paper, but it's not the 'best' GBD pipeline run.  What if you write multiple papers?  
- Do you want to keep track of all those by hand?  I don't!

1. Mark folders as 'keep', with a symlink and log entry that tells you it's important.
    1. You get all the same central log benefits as the 'best' symlink.
    

And your **Project Officer** gets things too!

1. Run reports on **all these logs simultaneously** about the status of all your important pipeline outputs.  



> That sounds pretty nice, but didn't you also say something about deleting folders?  Why do I need help doing that?


Ah!  I'm glad you asked.  The SymLink Tool will also:

1. Let you mark folders you want to 'remove' **before** you actually delete them.
    1. This is a **staging area** where you can take time to decide before you actually delete them.
    1. You can also 'unmark' them for removal if you change your mind.
1. When you want to really delete them you'll get:
    1. **Safety** 
        - The Symlink Tool will only delete folders that marked to 'remove'.
    1. **Provenance** 
        - You'll get a record in the central log telling you which pipeline runs were deleted, when, and you can add a note about why they were deleted.



> I'm still reading, and curious to see how this works.





--------------------------------------------------------------------------------



# Demonstration 


```{r}

```

